The \frb\ Campaign: Multi-Observatory Radio Burst and Implications for the FRB Population

### 2017-03-23 repeater-vladata.tex
working on volumetric rate and slsn connection

The burst rate analysis presented in \S \ref{sec:temp} shows that the detection probability of \frb\ is not stationary. As mentioned above, this could be caused by intrinsic events (e.g., magnetar outburst) or propagation effects. While a detailed analysis of temporal correlation will require more bursts and/or more sophistication, it is clear that \frb\ has significant correlation on short timescales \citep[sometimes called a ``red spectrum'';][]{2016MNRAS.458L..89C}. Another way of stating this is that the detection of a burst improves odds of finding more bursts in near-future observations.

If this statistical property describes other FRBs, it weakens previous constraints on repetition \citep{2015MNRAS.454..457P,2015ApJ...807...16L}. It is possible that other FRBs have repeating bursts, although calculating a limit requires knowing the temporal correlation function that defines the repetition.  Repetition also implies that there are fewer phyiscal sources of FRBs than implied by the burst detection rate.

FRBs like SGRs? 2016ApJ...826..226K

\subsection{FRB Rate}
% cut from gemini paper. TODO: incorporate referee comments on rate scale. Sarah?
{\color{red} Sarah: can you update this as a per-SFR rate?}
Assuming that \frb\ is from the same population as the other FRBs, the measurement of a distance for the host of \frb\ allows us to re-evaluate the cosmic volume and event rate per galaxy for the FRB population. The estimated projected FRB rate $R_p$\ understates the true rate by a beaming fraction $\Omega_b$\ \citep[$\sim$10\%;][]{1998MNRAS.298..625T}. For a comoving volume $V(z)$\ and galaxy number density $\Phi(M)$, the rate per galaxy is $R_{FRB} = R_p /(\Omega_b \Phi(M)V (z))$.

The first such rate estimate was made by \citet{2013Sci...341...53T}, who used the measured DM to estimate a characteristic distance for their sample of 4 FRBs. They assumed that all of the extragalactic component of the DM was caused by the IGM and scaled as $DM\approx z\times1200 \rm{pc}\ \rm{cm}^{-3}$\ \citep{2003ApJ...598L..79I,2004MNRAS.348..999I}. They also calculated the number of galaxies by assuming a characteristic $L_*$\ galaxy \citep[corresponding to stellar mass $M_* \approx 1011 M$;][]{2012MNRAS.421..621B}.

**add error bar?**
Our calculation differs in that we have better estimates of all three parameters. First, the projected FRB rate is now believed to be closer to $2\times10^3\ \rm{sky}^{-1} \rm{day}^{-1}$\ at high Galactic latitudes and flux densities brighter than 1 Jy ms \citep{2016arXiv161100458L,2016MNRAS.460L..30C, 2016MNRAS.455.2207R}. \frb\ is associated with a relatively small galaxy ($5\times10^7$\ Msun), which are roughly a factor of 100 more abundant \citep[$\Phi(M) \approx 10^{-2} \rm{Mpc}^{-3}$;][]{2007ApJ...665..265F}. Finally, the measured distance for FRB 121102 suggests that the extragalactic DM has comparable contributions from the IGM and host galaxy.

This reduces the characteristic distance by a factor of two and the volume by an order of magnitude. Considering these factors, we assume a characteristic redshift of the population that is twice that of \frb\ to estimate $R_{FRB} \approx 10^{-4} (0.1/\Omega_b) \rm{galaxy}^{-1} \rm{year}^{-1}$, which is two orders of magnitude lower than previously estimated \citep[assuming isotropic radiation;]{2013Sci...341...53T}.

There are significant caveats to the comparison of this rate to the rates of other classes of transient. This isotropic FRB rate assumes a projected FRB rate at a single (observationally defined) flux limit and that no bursts repeat. Using this calculation with data from more sensitive telescopes, we might infer a higher rate. However, if we assume that \frb\ is representative of the overall FRB population, then more sensitive observations would be more likely to find repeating FRBs that would effectively depress the estimate of the underlying isotropic rate per source. In this model, these two effects partially cancel, although without a physical model it is difficult to say how well.

\subsection{Observing Strategies}
As discussed above, \citet{2016MNRAS.458L..89C} note that the FRB repetition implies that the number of FRB-generating sources is smaller than implied by the number of bursts. In some scenarios, this means that there are areas on the sky that may contain no FRBs. Wide, shallow surveys are the preferred strategy for blind detection of FRBs. However, future efforts to localize FRBs are still most likely to succeed by targeting known FRBs to catch repetitions, assuming that all FRBs repeat.

>>>>

The chance of detecting a burst with the VLA changes substantially on day--month timescales. A major outstanding question is whether this time-variable rate is driven by an intrinsic process \citep{2016ApJ...826..226K} or extrinsic (propagation) effects \citep{CORDES}. However, the temporal distribution of VLA detections alone shows that \frb\ has significant correlation on short timescales. That is, detecting a burst predicts you are more likely to find another burst soon thereafter and not detecting a burst predicts you will likey not detect another burst soon thereafter. While this is sometimes referred to as a ''red spectrum'' \citep{2016MNRAS.458L..89C}, the observations presented here are only sensitive to timescales of days (time between observations) and months (time between campaigns).

If this statistical property describes other FRBs, it weakens previous constraints on repetition \citep{2015MNRAS.454..457P,2015ApJ...807...16L}. It is possible that other FRBs have repeating bursts, but calculating a limit requires knowing their temporal correlation function. Wide, shallow surveys are the preferred strategy for blind detection of FRBs. Also, since ''bursts predict bursts'', many short observations are more likely to blindly detect an FRB than a single long observation of the same total length.

\subsection{Rate Comparison}

The calculation of a volumetric rate has been used to identify classes of transient and their defining physical processes \citep{2006ARA&A..44..507W, 2012Sci...337..927G}. \citet{2013Sci...341...53T} used this approach to estimate an isotropic rate of $10^{-3}$\ galaxy$^{-1}$\ yr$^{-1}$ by assuming a characteristic redshift horizon of 0.9. This rate is high \citep[comparable to the rate of core-collapse supernovae;][]{2006Natur.439...45D}, which has been used to argue that FRBs are likely not associated with other classes of transient, such as GRBs or rarer SNe.

This conclusion is in tension with new models inspired by multi-wavelength associations to \frb. Its coincidence with a compact, persistent radio source is consistent with \frb as a young magnetar in a luminous pulsar wind nebula \citep{2017arXiv170104815K}. At the same time, \frb is found in a low-metallicity dwarf galaxy, which is the preferred environment for long GRBs and Type-I superluminous supernovae \citep[LGRB and SLSN-I, respectively;][]{2017arXiv170102370M}.

We now reevaluate this rate using the properties of \frb\ as a prototype FRB. First, we recast this estimate as a beamed, volumetric rate, $R_{FRB} = R_p /(\Omega_b V (z))$, where $R_p$\ is the projected FRB rate, $\Omega_b$\ is the beaming fraction \citep[$\sim0.1$;][]{1998MNRAS.298..625T} and $V(z)$\ is the volume out to redshift $z$. Second, the projected FRB rate is now believed to be lower than initially believed \citep[$2\times10^3\ \rm{sky}^{-1} \rm{day}^{-1}$\ at high Galactic latitudes and flux densities brighter than 1 Jy ms; ][]{2016arXiv161100458L, 2016MNRAS.460L..30C, 2016MNRAS.455.2207R}. Including these effects, we estimate a volumetric rate of $6\times10^{-5} (0.1/\Omega_b)$\ Mpc$^{-3}$\ yr$^{-1}$, assuming a typical FRB horizon redshift of 0.9. For comparison, the LGRB and SLSN-I rates are orders of magnitude lower, near $10^{-8}$ Mpc$^{-3}$\ yr$^{-1}$ \citep{2012Sci...337..927G}.

The revised FRB rate is an improvement, but it still ignores multiple effects. The burst spectral structure discussed here shows that this estimate likely underestimates the FRB rate by missing bursts that fall outside the observing band (typically a few hundred MHz near 1.4~GHz). For a typical burst envelope width of 500~MHz centered at GHz-scale frequencies, this could underestimate the rate by a factor of a few. The horizon scale ($z=0.9$) is not well defined and can change the implied rate by an order of magnitude.

sensitivity is the biggest awkward factor!
an out: use luminosity distribution for absolute rate scale?

calculate number of repetitions and compare to Metzger limit

%DM dominated by the IGM \citep[$DM\approx z\times1200 \rm{pc}\ \rm{cm}^{-3}$ for fully ionized hydrogen or $DM\approx z\times900 \rm{pc}\ \rm{cm}^{-3}$ for more realistic ionization state;][]{2003ApJ...598L..79I,2004MNRAS.348..999I}. 
%galaxy number density $\Phi(M)$
%They also calculated the number of galaxies by assuming a characteristic $L_*$\ galaxy \citep[corresponding to stellar mass $M_* \approx 1011 M$;][]{2012MNRAS.421..621B}.
%
%This isotropic FRB rate assumes a projected FRB rate at a single (observationally defined) flux limit and that no bursts repeat. Using this calculation with data from more sensitive telescopes, we might infer a higher rate. However, if we assume that \frb\ is representative of the overall FRB population, then more sensitive observations would be more likely to find repeating FRBs that would effectively depress the estimate of the underlying isotropic rate per source. In this model, these two effects partially cancel, although without a physical model it is difficult to say how well.
